---
title: "2. Time Series Graphics.RMD"
author: "Russ Conte"
date: "10/2/2021"
output: html_document
---

# 2. Time Series Graphics

## 2.1 `tsibble` objects

How to create a `tsibble`:

```{r 2.1 Creating a tsibble}
library(fpp3)
y <- tsibble(
  Year = 2015:2019,
  Observation  = c(123, 39, 78, 52, 110),
  index = Year
)

y
```

For observations that are more frequent than once per year, we need to use a time class function on the index. For example:

```{r Key bariables}
olympic_running

```

The 14 time series in this object are uniquely identified by the keys: the `Length` and `Sex` variables. The `distinct()` function can be used to show the categories of each variable or even combinations of variables:

```{r Using the distinct function to show categories of each variable}
olympic_running %>% distinct(Sex)
```

## Working with tsibble objects

We can use dplyr functions such as `mutate()`, `filter()`, `select()` and `summarise()` to work with `tsibble` objects.

```{r Look at the PBS data set}
PBS
```

Filter the dataset to extract the A10 scripts:

```{r Filter the dataset to extract the A10 scripts}
a10 <- PBS %>% filter(ATC2 == "A10")
a10
```

Select the columns we need in our analysis:

```{r Select the columns we need in our analysis}
a10 <- a10 %>% select(Month, Concession, Type, Cost)
a10

```

Summarise allows us to combine data across keys:

```{r example of summarise}
a10 <- a10 %>% summarise(TotalCost = sum(Cost))
a10
```

Create new variables using the `mutate` function:

```{r}
a10 <- a10 %>% mutate(TotalCost = TotalCost / 1e6)
a10
```

## Read a CSV file and convert to a table:

```{r Read a CSV file and convert to a table}
prison <- readr::read_csv('https://OTexts.com/fpp3/extrafiles/prison_population.csv')

prison <- prison %>% 
  mutate(Quarter = yearquarter(Date)) %>% 
  select(-Date) %>% 
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)
prison

```

## 2.2 Time plots

For time series data, the obvious graph to start with is a time plot. That is, the observations are plotted against the time of observation, with consecutive observations joined by straight lines. Figure 2.1 shows the weekly economy passenger load on Ansett airlines between Australia’s two largest cities.

```{r Time series plot}
melsyd_economy <- ansett %>% 
  filter(Airports == "MEL-SYD", Class == "Economy") %>% 
  mutate(Passengers = Passengers / 1000)
autoplot(melsyd_economy, Passengers) +
  labs(title = "Ansett airline economy class",
       subtitle = "Melbourne-Sydney",
       y = "Passengers ('000)")
```

The time plot immediately reveals some interesting features.

• There was a period in 1989 when no passengers were carried — this was due to an industrial dispute.<br>
• There was a period of reduced load in 1992. This was due to a trial in which some economy class seats were replaced by business class seats.<br>
• A large increase in passenger load occurred in the second half of 1991.<br>
• There are some large dips in load around the start of each year. These are due to holiday effects.
• There is a long-term fluctuation in the level of the series which increases during 1987, decreases in 1989, and increases again through 1990 and 1991.<br>
• There are some periods of missing observations.<br>

Let's take a look at the a10 data saved earlier, plotted as a time series:

```{r a10 data saved as a time series}
autoplot(a10, TotalCost) +
  labs(y = "$ millionss",
       title = "Australian antidiabetic drug sales")

```

## 2.3 Time Series Patterns

<span style="color:#0000ff;">Trend</span><br>
A trend exists when there is a long-term increase or decrease in the data. It does not have to be linear. Sometimes we will refer to a trend as “changing direction,” when it might go from an increasing trend to a decreasing trend. There is a trend in the antidiabetic drug sales data shown in Figure 2.2.<br>
<span style="color:#0000ff;">Seasonal</span><br>
A seasonal pattern occurs when a time series is affected by seasonal factors such as the time of the year or the day of the week. Seasonality is always of a fixed and known period. The monthly sales of antidiabetic drugs (Figure 2.2) shows seasonality which is induced partly by the change in the cost of the drugs at the end of the calendar year.<br>
<span style="color:#0000ff;">Cyclic</span><br>
A cycle occurs when the data exhibit rises and falls that are not of a fixed frequency. These fluctuations are usually due to economic conditions, and are often related to the “business cycle.” The duration of these fluctuations is usually at least 2 years.<br>

## 2.4 Seasonal plots

A seasonal plot is similar to a time plot except that the data are plotted against the individual “seasons” in which the data were observed. An example is given in Figure 2.4 showing the antidiabetic drug sales.

```{r Seasonal plot: Antidiabetic drug sales}
a10 %>%
  gg_season(TotalCost, labels = "both") +
  labs(y = "$ millions",
       title = "Seasonal plot: Antidiabets drug sales") +
  expand_limits(x = ymd(c("1972-12-28", "1973-12-04")))
```

### Multiple seasonal periods

Where the data has more than one seasonal pattern, the period argument can be used to select which seasonal plot is required. The vic_elec data contains half-hourly electricity demand for the state of Victoria, Australia. We can plot the daily pattern, weekly pattern or yearly pattern by specifying the period argument.

```{r Plotting multiple seasonal periods}
vic_elec %>% gg_season(Demand, period = "day") +
  theme(legend.position = "none") +
  labs(y = "MW", title = "Electricity demand: Victoria")

```



```{r Seasonal plot showing weekly seasonal patterns for Victorian electricity demand.}

vic_elec %>% gg_season(Demand, period = "week") +
  theme(legend.position = "none") +
  labs(y = "MW", title = "Electricity demand: Victoria by day of the week")

```

```{r Seasonal plot showing yearly seasonal patterns for Victorian electricity demand.}

vic_elec %>% gg_season(Demand, period = "year") +
  labs(y="Megawatts", title = "Electricity demand: Victoria") 

```

## 2.5 Seasonal subseries plots

An alternative plot that emphasises the seasonal patterns is where the data for each season are collected together in separate mini time plots.

```{r Seasonal subseries plot of monthly antidiabetic drug sales in Australia}
a10 %>% 
  gg_subseries(TotalCost) +
  labs(
    y = "$ millions",
    title = "Australian antidiabetic drug sales"
  )

```

### Example: Australian holiday tourism

```{r Plotting holiday tourism}

holidays <- tourism %>% 
  filter(Purpose == "Holiday") %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))

holidays

```

```{r Australian domestic holidays by state}
autoplot(holidays, Trips) +
  labs(y = "Overnight trips '000",
       title = "Australian domestic holidays") +
  facet_grid(~State)

```

To see the timing of the seasonal peaks in each state, we can use a season plot.

```{r Seasonal plot example}
gg_season(holidays, Trips) +
  labs(y = "Overnight trips '000",
       title = "Australian domestic holidays")

```

```{r Subseries plots of Australian domestic holidays by state}
holidays %>%
  gg_subseries(Trips) +
  labs(y = "Overnight trips '000",
       titla = "Australian domestic holidays")

```

## 2.6 Scatterplots

The graphs discussed so far are useful for visualising individual time series. It is also useful to explore relationships between time series.

```{r Half-hour electricity demand in Victoria}
vic_elec %>% 
  filter(year(Time) == 2014) %>% 
  autoplot(Demand)+
  labs(y = "Gigawatts",
       title = "Half-hourly electricity demand: Victoria")
```

```{r Half-hour temperatures: Melbourne, Australia}
vic_elec %>% 
  filter(year(Time) == 2014) %>% 
  autoplot(Temperature) +
  labs(
    y = "Degrees Celsius",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```

```{r Relationship between electricity demand and temperature for Victoria in 2014}
vic_elec %>% 
  filter(year(Time) == 2014) %>% 
  ggplot(aes(x = Temperature, y = Demand)) +
  geom_point() +
  labs(x = "Temperature (degrees Celsius)",
    title = "Electricity demand (gigawatts)")
```

### Correlation

$$r = \frac{\sum_{i=0}^n\left(x_t - \bar{x} \right)\left(y_t - \bar{y} \right)^2}{\sqrt{\left(x_t - \bar{x} \right)^2}\sqrt{\left(y_t - \bar{y} \right)^2}}$$

### Scatterplot matrices

When there are several potential predictor variables, it is useful to plot each variable against each other variable.

```{r Quarterly visitor nights for the states and territories of Australia.}
visitors <- tourism %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))
visitors %>% 
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State), scales = "free_y") +
  labs(title = "Australian domestic tourism",
       y = "Overnight trips '000")

```


To see the relationships between these eight time series, we can plot each time series against the others. These plots can be arranged in a scatterplot matrix, as shown in Figure 2.18. (This plot requires the GGally package to be installed.)

```{r View correlations of time seris}
visitors %>% 
  pivot_wider(values_from = Trips, names_from = State) %>% 
  GGally::ggpairs(columns = 2:9)

```

For each panel, the variable on the vertical axis is given by the variable name in that row, and the variable on the horizontal axis is given by the variable name in that column. There are many options available to produce different plots within each panel. In the default version, the correlations are shown in the upper right half of the plot, while the scatterplots are shown in the lower half. On the diagonal are shown density plots.

The value of the scatterplot matrix is that it enables a quick view of the relationships between all pairs of variables. In this example, mostly positive relationships are revealed, with the strongest relationships being between the neighbouring states located in the south and south east coast of Australia, namely, New South Wales, Victoria and South Australia. Some negative relationships are also revealed between the Northern Territory and other regions. The Northern Territory is located in the north of Australia famous for its outback desert landscapes visited mostly in winter. Hence, the peak visitation in the Northern Territory is in the July (winter) quarter in contrast to January (summer) quarter for the rest of the regions.

## 2.7 Lab plots

Figure 2.19 displays scatterplots of quarterly Australian beer production (introduced in Figure 1.1), where the horizontal axis shows lagged values of the time series. Each graph shows  $$y_t \textrm{ is plotted against } y_{t-k} \textrm{ for different values of k}$$
```{r Lagged scatterplots for quarterly beer production.}

recent_production <- aus_production %>% 
  filter(year(Quarter) >= 2000)
recent_production %>% 
  gg_lag(Beer, geom = "point") +
  labs(x = "lab(Beerk, k")

```

Here the colours indicate the quarter of the variable on the vertical axis. The relationship is strongly positive at lags 4 and 8, reflecting the strong seasonality in the data. The negative relationship seen for lags 2 and 6 occurs because peaks (in Q4) are plotted against troughs (in Q2)