---
title: "7. Time_Series_Regression_Models.RMD"
author: "Russ Conte"
date: "10/17/2021"
output: html_document
---

In this chapter we discuss regression models. The basic concept is that we forecast the time series of interest y assuming that it has a linear relationship with other time series $x$.

For example, we might wish to forecast monthly sales $y$ using total advertising spend $x$ as a predictor. Or we might forecast daily electricity demand y using temperature $x_1$ and the day of week $x_2$ as predictors.

The <b>forecast variable</b> $y$ is sometimes also called the regressand, dependent or explained variable. The <b>predictor variables</b> $x$ are sometimes also called the regressors, independent or explanatory variables. <mark><b>In this book we will always refer to them as the “forecast” variable and “predictor” variables.</b></mark>

## 7.1 The Linear Model

In the simplest case, the regression model allows for a linear relationship between the forecast variable y and a single predictor variable $x$:

$$y_t = \beta_0 + \beta_1x_t+\epsilon_t$$

An artificial example of data from such a model is shown in Figure 7.1. The coefficients $\beta_0$ and $\beta_1$ denote the intercept and the slope of the line respectively. The intercept $\beta_0$ represents the predicted value of $y$ when $x$=0. The slope $\beta_1$ represents the average predicted change in $y$ resulting from a one unit increase in $x$.

```{r}
library(tidyverse)
library(fpp3)
ggplot(data = mtcars,mapping = aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle(label = 'Wt vs mpg for MTCars')

```

## Example: US Consumption Expenditure

Figure 7.2 shows time series of quarterly percentage changes (growth rates) of real personal consumption expenditure, $y$, and real personal disposable income, $x$, for the US from 1970 Q1 to 2019 Q2.

```{r Figure 7.2: Percentage changes in personal consumption expenditure and personal income for the US.}

us_change %>% 
  pivot_longer(c(Consumption, Income), names_to = "Series") %>% 
  autoplot(value) +
  labs(y = "% change")

```

A scatter plot of the consumption changes against income changes is showin in Figure 7.3, along with the estimated regression line:

$$\hat{y_t} = 0.54 + 0.27x_t$$

```{r Figure 7.3: Scatterplot of quarterly changes in consumption expenditure versus quarterly changes in personal income and the fitted regression line.}

us_change %>% 
  ggplot(aes(x = Income, y = Consumption)) +
  labs(y = "Consumption (quarterly % change)",
       x = "Income (quarterly % change)") +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)

```

The equation is estimated using the `TSLM()` (time series liner model) function:

```{r Time series linear model applied to US_change}
us_change %>% 
  model(TSLM(Consumption ~ Income)) %>% 
  report()
```

## Multiple Linear Regression

When there are two or more predictor variables, the model is called a multiple regression model. The general form of a multiple regression model is

$$y_t=\beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t} + ... + \beta_k x_{k, t} + \epsilon_t$$

where $y$ is the variable to be forecast and $x_1,…,x_k$ are the $k$ predictor variables. Each of the predictor variables must be numerical. The coefficients $\beta_1,...\beta_k$ measure the effect of each predictor after taking into account the effects of all the other predictors in the model. Thus, the coefficients measure the marginal effects of the predictor variables.

## Example: US Consumption Expenditure

Figure 7.4 shows additional predictors that may be useful for forecasting US consumption expenditure. These are quarterly percentage changes in industrial production and personal savings, and quarterly changes in the unemployment rate (as this is already a percentage). Building a multiple linear regression model can potentially generate more accurate forecasts as we expect consumption expenditure to not only depend on personal income but on other predictors as well.

```{r}
# Answered at: https://stackoverflow.com/questions/69606489/plotting-each-variable-in-a-time-series-on-the-same-plot/69606629#69606629
library(tidyverse)
us_change %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(Quarter, value, color = name)) +
  geom_line() +
  facet_wrap(~name, ncol = 1, scales = "free_y") +
  guides(color = "none")


```

Figure 7.5 is a scatterplot matrix of five variables. The first column shows the relationships between the forecast variable (consumption) and each of the predictors. The scatterplots show positive relationships with income and industrial production, and negative relationships with savings and unemployment. The strength of these relationships are shown by the correlation coefficients across the first row. The remaining scatterplots and correlation coefficients show the relationships between the predictors.

```{r Figure 7.5: A scatterplot matrix of US consumption expenditure and the four predictors. }

us_change %>% 
  GGally::ggpairs(columns = 2:6)

```

## Assumptions

When we use a linear regression model, we are implicitly making some assumptions about the variables in Equation (7.1).

First, we assume that the model is a reasonable approximation to reality; that is, the relationship between the forecast variable and the predictor variables satisfies this linear equation.

Second, we make the following assumptions about the errors $( \epsilon_1...\epsilon_T)$

•   they have mean zero; otherwise the forecasts will be systematically biased.
•   they are not autocorrelated; otherwise the forecasts will be inefficient, as there is more information in the data that can be exploited.
•    they are unrelated to the predictor variables; otherwise there would be more information that should be included in the systematic part of the model.

It is also useful to have the errors being normally distributed with a constant variance $\sigma^2$

in order to easily produce prediction intervals.

Another important assumption in the linear regression model is that each predictor $x$
is not a random variable. If we were performing a controlled experiment in a laboratory, we could control the values of each $x$ (so they would not be random) and observe the resulting values of $y$. With observational data (including most data in business and economics), it is not possible to control the value of $x$, we simply observe it. Hence we make this an assumption.

## 7.2 Least Squares Estimation

In practice, of course, we have a collection of observations but we do not know the values of the coefficients $\beta_0, \beta_1, ... \beta_k$. These need to be estimated from the data.

The least squares principle provides a way of choosing the coefficients effectively by minimising the sum of the squared errors. That is, we choose the values of $\beta_0, \beta_1, ... \beta_k$ that minimise

$\sum_{t=1}^T \epsilon_t^2 = \sum_{t=1}^T (y_t - \beta_0 - \beta_1 x_{1,t} - \beta_2 x_{2,t} - ... - \beta_k x_{k, t})^2$

This is called <b>least squares</b> estimation because it gives the least value for the sum of squared errors. Finding the best estimates of the coefficients is often called “fitting” the model to the data, or sometimes “learning” or “training” the model. The line shown in Figure 7.3 was obtained in this way.

The `TSLM()` function fits a linear regression model to time series data. It is similar to the `lm()` function which is widely used for linear models, but `TSLM()` provides additional facilities for handling time series.

## Example: US Consumption Expenditure

A multiple linear regression model for US consumption is:

$y_t = \beta_0 + \beta_1 x_{1,t} + \beta_2 x_{2,t}  + \beta_3 x_{3,t} + \beta_4 x_{4,t} +  \epsilon_t$

where $y$ is the percentage change in real personal consumption expenditure, $x_1$ is the percentage change in real personal disposable income, $x_2$ is the percentage change in industrial production, $x_3$ is the percentage change in personal savings and $x_4$ is the change in the unemployment rate.

The following output provides information about the fitted model. The first column of `Coefficients` gives an estimate of each $\beta$ coefficient and the second column gives its standard error (i.e., the standard deviation which would be obtained from repeatedly estimating the $\beta$ coefficients on similar data sets). The standard error gives a measure of the uncertainty in the estimated $\beta$ coefficient.

```{r Information about the fitted model}

fit_consMR <- us_change %>% 
  model(tslm = TSLM((Consumption ~ Income + Production + Unemployment + Savings)))

report(fit_consMR)

```

For forecasting purposes, the final two columns are of limited interest. The “t value” is the ratio of an estimated $\beta$ coefficient to its standard error and the last column gives the p-value: the probability of the estimated $\beta$ coefficient being as large as it is if there was no real relationship between consumption and the corresponding predictor. <b>This is useful when studying the effect of each predictor, but is not particularly useful for forecasting.</b>

## Fitted Values

Predictions of y can be obtained by using the estimated coefficients in the regression equation and setting the error term to zero. In general we write:

$$\hat{y} = \hat{\beta_0} + \hat{\beta_1 x_{1,t}} +\hat{\beta_2}x_{2,t} + ... + \hat{\beta_k} x_{k,t} $$
Plugging in the values of $x_{1,t}, ... ,x_{k,t}$ for $t$=1,…,T returns predictions of $y_t$ within the training set, referred to as fitted values. Note that these are predictions of the data used to estimate the model, not genuine forecasts of future values of $y$.

The following plots show the actual values compared to the fitted values for the percentage change in the US consumption expenditure series. The time plot in Figure 7.6 shows that the fitted values follow the actual data fairly closely. This is verified by the strong positive relationship shown by the scatterplot in Figure 7.7.

```{r Figure 7.6: Time plot of actual US consumption expenditure and predicted US consumption expenditure.}

augment(fit_consMR) %>% 
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Consumption, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
       title = "Percentage change in US consumption expenditure") +
  scale_color_manual(values = c(Data = "black", Fitted = "#D55E00")) +
  guides(colour = guide_legend(title = "Legend"))

```

```{r Figure 7.7: Actual US consumption expenditure plotted against predicted US consumption expenditure.}

augment(fit_consMR) %>% 
  ggplot(aes(x = Consumption, y = .fitted)) +
  geom_point() +
  labs(
    y = "Fitted (predicted values)",
    x = "Data (actual values)",
    title = "Percent change in US consumption expendutire"
  ) +
  geom_abline(intercept = 0, slope = 1)

```

## Goodness-of-Fit

A common way to summarise how well a linear regression model fits the data is via the coefficient of determination, or $R^2$. This can be calculated as the square of the correlation between the observed $y$ values and the predicted $\hat{y}$ values. Alternatively, it can also be calculated as,

$$R^2 = \frac{\sum(\hat{y}_{t} - \bar{y})^2}{\sum(y_{t}-\bar{y})^2},$$

where the summations are over all observations. Thus, it reflects the proportion of variation in the forecast variable that is accounted for (or explained) by the regression model.

In simple linear regression, the value of $R^2$is also equal to the square of the correlation between 
$y$ and $X$ (provided an intercept has been included).

If the predictions are close to the actual values, we would expect $R^2$ to be close to 1. On the other hand, if the predictions are unrelated to the actual values, then $R^2$=0 (again, assuming there is an intercept). In all cases, $R^2$ lies between 0 and 1.

The $R^2$ value is used frequently, though often incorrectly, in forecasting. The value of $R^2$ will never decrease when adding an extra predictor to the model and this can lead to over-fitting. There are no set rules for what is a good $R^2$ value, and typical values of $R^2$ depend on the type of data used. Validating a model’s forecasting performance on the test data is much better than measuring the $R^2$ value on the training data.

Figure 7.7 plots the actual consumption expenditure values versus the fitted values. The correlation between these variables is $r$=0.877 hence $R^2$=0.768 (shown in the output above). In this case model does an excellent job as it explains 76.8% of the variation in the consumption data. Compare that to the $R^2$ value of 0.15 obtained from the simple regression with the same data set in Section 7.1. Adding the three extra predictors has allowed a lot more of the variation in the consumption data to be explained.

## Standard Error of the Regression

Another measure of how well the model has fitted the data is the standard deviation of the residuals, which is often known as the “residual standard error.” This is shown in the above output with the value 0.31.

It is calculated using:

$\hat{\sigma_e} = \sqrt {\frac{1}{T - k - 1}\sum_{t = 1}^T e_t^2}$

where k is the number of predictors in the model. Notice that we divide by T−k−1 because we have estimated k+1 parameters (the intercept and a coefficient for each predictor variable) in computing the residuals.

The standard error is related to the size of the average error that the model produces. We can compare this error to the sample mean of y or with the standard deviation of y

to gain some perspective on the accuracy of the model.

The standard error will be used when generating prediction intervals, discussed in Section 7.6.

# 7.3 Evaulating the regression model

The differences between the observed y values and the corresponding fitted ^y values are the training-set errors or “residuals” defined as:

$ e_t = y_t - \hat{y_t}$
$\text{   } = y_t - \hat{\beta_1{x_{1,t}}} - \hat{\beta_2{x_{2,t}}} - ... - \hat{\beta_k{x_{k,t}}}$

for $t = 1,...T.$ Each residual is the unpredictable component of the associated observation.

As a result of these properties, it is clear that the average of the residuals is zero, and that the correlation between the residuals and the observations for the predictor variable is also zero. (This is not necessarily true when the intercept is omitted from the model.)

## ACF plot of residuclas

With time series data, it is highly likely that the value of a variable observed in the current time period will be similar to its value in the previous period, or even the period before that, and so on. Therefore when fitting a regression model to time series data, it is common to find autocorrelation in the residuals. In this case, the estimated model violates the assumption of no autocorrelation in the errors, and our forecasts may be inefficient — there is some information left over which should be accounted for in the model in order to obtain better forecasts. The forecasts from a model with autocorrelated errors are still unbiased, and so are not “wrong,” but they will usually have larger prediction intervals than they need to. Therefore we should always look at an ACF plot of the residuals.

## Histogram of residuals

It is always a good idea to check whether the residuals are normally distributed. As we explained earlier, this is not essential for forecasting, but it does make the calculation of prediction intervals much easier.

### Example
using the `gg_tsresiduals()` function introduced in Section 5.3, we can obtain all the useful residual diagnostics mentioned above:

```{r  Figure 7.8: Analysing the residuals from a regression model for US quarterly consumption.}

fit_consMR %>% gg_tsresiduals()

```

```{r}

augment(fit_consMR) %>% 
  features(.innov, ljung_box, lag = 10, dof = 5)
```

The time plot shows some changing variation over time, but is otherwise relatively unremarkable. This heteroscedasticity will potentially make the prediction interval coverage inaccurate.

The histogram shows that the residuals seem to be slightly skewed, which may also affect the coverage probability of the prediction intervals.

## Residual plot against predictors

We would expect the residuals to be randomly scattered without showing any systematic patterns. A simple and quick way to check this is to examine scatterplots of the residuals against each of the predictor variables. If these scatterplots show a pattern, then the relationship may be nonlinear and the model will need to be modified accordingly. See Section 7.7 for a discussion of nonlinear regression

### Example

The residuals from the multiple regression model for forecasting US consumption plotted against each predictor in Figure 7.9 seem to be randomly scattered. Therefore we are satisfied with these in this case.

```{r Figure 7.9: Scatterplots of residuals versus each predictor. }

us_change %>% 
  left_join(residuals(fit_consMR), by = "Quarter") %>% 
  pivot_longer(Income:Unemployment, names_to = "regressor", values_to = "x") %>% 
  ggplot(aes(x = x, y = .resid)) +
  geom_point()+
  facet_wrap(.~ regressor, scales = "free_x") +
  labs(y = "Residuals", x = "")
```

## Residuals against fitted values

A plot of the residuals against the fitted values should also show no pattern. If a pattern is observed, there may be “heteroscedasticity” in the errors which means that the variance of the residuals may not be constant. If this problem occurs, a transformation of the forecast variable such as a logarithm or square root may be required (see Section 3.1).

### Example

Continuing the previous example, Figure 7.10 shows the residuals plotted against the fitted values. The random scatter suggests the errors are homoscedastic.

```{r Figure 7.10: Scatterplots of residuals versus fitted values.}
augment(fit_consMR) %>% 
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() +
  labs(x = "Fitted", y = "Residuals")

```

## Spurious regression

More often than not, time series data are “non-stationary”; that is, the values of the time series do not fluctuate around a constant mean or with a constant variance. We will deal with time series stationarity in more detail in Chapter 9, but here we need to address the effect that non-stationary data can have on regression models.

Regressing non-stationary time series can lead to spurious regressions. The output of regressing Australian air passengers on rice production in Guinea is shown in Figure 7.13. High R2

and high residual autocorrelation can be signs of spurious regression. Notice these features in the output below. We discuss the issues surrounding non-stationary data and spurious regressions in more detail in Chapter 10.

Cases of spurious regression might appear to give reasonable short-term forecasts, but they will generally not continue to work into the future.

```{r Figure 7.13: Residuals from a spurious regression.}

fit <- aus_airpassengers %>% 
  filter(Year <= 2011) %>% 
  left_join(guinea_rice,by = "Year") %>% 
  model(TSLM(Passengers ~ Production))
report(fit)
```

Conclusion through section 7.3

## 7.4 Some useful predictors

### Trend

It is common for time series data to be trending. A linear trend can be modelled by simply using $x_{1,t} = t$ as a predictor,
$$y_t = \beta_0 + \beta_1{t} + \epsilon_t$$
where $t = 1,...T.$ A trend variable can be specified in the `TSLM()` function using the `trend()` special. In section 7.7 we discuss how we can also model nonlinear trends

### Dummy variables

So far, we have assumed that each predictor takes numerical values. But what about when a predictor is a categorical variable taking only two values (e.g., “yes” and “no”)? Such a variable might arise, for example, when forecasting daily sales and you want to take account of whether the day is a public holiday or not. So the predictor takes value “yes” on a public holiday, and “no” otherwise.

This situation can still be handled within the framework of multiple regression models by creating a “dummy variable” which takes value 1 corresponding to “yes” and 0 corresponding to “no.” A dummy variable is also known as an “indicator variable.”

A dummy variable can also be used to account for an outlier in the data. Rather than omit the outlier, a dummy variable removes its effect. In this case, the dummy variable takes value 1 for that observation and 0 everywhere else. An example is the case where a special event has occurred. For example when forecasting tourist arrivals to Brazil, we will need to account for the effect of the Rio de Janeiro summer Olympics in 2016.

If there are more than two categories, then the variable can be coded using several dummy variables (one fewer than the total number of categories). TSLM() will automatically handle this case if you specify a factor variable as a predictor. There is usually no need to manually create the corresponding dummy variables.

### Example: Australian quarterly beer production

```{r Figure 7.14: Australian quarterly beer production.}
recent_production <- aus_production %>% 
  filter(year(Quarter) >= 1992)
recent_production %>% 
  autoplot(Beer) +
  labs(y = "Megalitres", title = "Australian quarterly beer production")

```

We want to forecast the value of future beer production. We can model this data using a regression model with a linear trend and quarterly dummy variables,

$y_t = \beta_0 + \beta_1{t} + \beta_2{d_{2,t}} + \beta_3{d_{3,t}} + \beta_4{d_{4,t}} + \epsilon_t$

where $d_{i,t}$ = 1, if $t$ is in quarter $i$, and $0$ otherwise. The first quarter variable has been omitted, so the coefficients associated with the other quarters are measures of the difference between those quarters and the first quarter.

```{r}

fit_beer <- recent_production %>% 
  model(TSLM(Beer ~ trend() + season()))
report(fit_beer)

```

Note that trend() and season() are not standard functions; they are “special” functions that work within the TSLM() model formulae.

There is an average downward trend of -0.34 megalitres per quarter. On average, the second quarter has production of 34.7 megalitres lower than the first quarter, the third quarter has production of 17.8 megalitres lower than the first quarter, and the fourth quarter has production of 72.8 megalitres higher than the first quarter.

```{r Figure 7.15: Time plot of beer production and predicted beer production.}

augment(fit_beer) %>% 
  ggplot(aes(x = Quarter)) +
  geom_line(aes(y = Beer, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  scale_color_manual(
    values = c(Data = "black", Fitted = "#D55E00")
  ) +
  labs(y = "Megaliters",
       title = "Austrailian quarterly beer production") +
  guides(colour = guide_legend(title = "Series"))

```

```{r}
augment(fit_beer) %>%
  ggplot(aes(x = Beer, y = .fitted,
             colour = factor(quarter(Quarter)))) +
  geom_point() +
  labs(y = "Fitted", x = "Actual values",
       title = "Australian quarterly beer production") +
  geom_abline(intercept = 0, slope = 1) +
  guides(colour = guide_legend(title = "Quarter"))

```

## Intervention variables

It is often necessary to model interventions that may have affected the variable to be forecast. For example, competitor activity, advertising expenditure, industrial action, and so on, can all have an effect.

When the effect lasts only for one period, we use a “spike” variable. This is a dummy variable that takes value one in the period of the intervention and zero elsewhere. A spike variable is equivalent to a dummy variable for handling an outlier.

## Distributed lags

It is often useful to include advertising expenditure as a predictor. However, since the effect of advertising can last beyond the actual campaign, we need to include lagged values of advertising expenditure. Thus, the following predictors may be used. 
$x_1$ = advertising for previous month;
$x_2$ =advertising for two months previously;
.
.
.
$x_m$ = advertising for $m% months previously.

$x_{1,t} = sin{\frac{2\pi t}{m}}, x_{2,t} = cos{\frac{2\pi t}{m}}, x_{3,t} = sin{\frac{4\pi t}{m}},$
$x_{4,t} = sin{\frac{4\pi t}{m}}, x_{5,t} = cos{\frac{6\pi t}{m}}, x_{6,t} = sin{\frac{6\pi t}{m}},$

and so on. If we have monthly seasonality, and we use the first 11 of these predictor variables, then we will get exactly the same forecasts as using 11 dummy variables.

With Fourier terms, we often need fewer predictors than with dummy variables, especially when m is large. This makes them useful for weekly data, for example, where m≈52. For short seasonal periods (e.g., quarterly data), there is little advantage in using Fourier terms over seasonal dummy variables.

These Fourier terms are produced using the fourier() function. For example, the Australian beer data can be modelled like this.

```{r}
fourier_beer <- recent_production %>% 
  model(TSLM(Beer ~ trend() + fourier(K = 2)))
report(fourier_beer)
```

The `K` argument to `fourier()` specifies how many pairs of sin and cos terms to include. The maximum allowed is $K=\frac{m}{2}$ where $m$is the seasonal period. Because we have used the maximum here, the results are identical to those obtained when using seasonal dummy variables.

If only the first two Fourier terms are used ($x_{1,t} \text{ and } x_{2,t}$), the seasonal pattern will follow a simple sine wave. A regression model containing Fourier terms is often called a harmonic regression because the successive Fourier terms represent harmonics of the first two Fourier terms.

Conclusion through section 7.4

## 7.5 Selecting predictors

In this section,  we will use a measure of predictive accuracy. Five such measures are introduced in this section. They can be shown using the glance() function, here applied to the model for US consumption:

```{r Five measures of predcitve accuracy}

glance(fit_consMR) %>% 
  select(adj_r_squared, CV, AIC, AICc, BIC)

```

We compare these values against the corresponding values from other models. For the CV, AIC, AICc and BIC measures, we want to find the model with the lowest value; for Adjusted R2, we seek the model with the highest value.

## Adjusted R^2^

Computer output for a regression will always give the R2 value, discussed in Section 7.2. However, it is not a good measure of the predictive ability of a model. It measures how well the model fits the historical data, but not how well the model will forecast future data.

An alternative which is designed to overcome these problems is the adjusted R^2^:

$\bar{R}^2 = 1 - (1 - R^2) \frac{T - 1}{T - k -1}$

where $T$ is the number of observations and $k$ is the number of predictors. This is an improvement on $R^2$, as it will no longer increase with each added predictor. Using this measure, the best model will be the one with the largest value of $\bar{R}^2$. Maximising $\bar{R}^2$ is equivalent to minimising the standard error $\hat{\sigma_e}$ given in Equation 7.3

Maximising $R^2$ works quite well as a method of selecting predictors, although it does tend to err on the side of selecting too many predictors.

## Cross-validation

Time series cross-validation was introduced in Section 5.8 as a general tool for determining the predictive ability of a model. For regression models, it is also possible to use classical leave-one-out cross-validation to selection predictors (Bergmeir et al., 2018). This is faster and makes more efficient use of the data. The procedure uses the following steps:

1. Remove observation $t$ from the data set, and fit the model using the remaining data. Then compute the error ($e_t^* = y_t - \hat{y_t}$) for the omitted observation. (This is not the same as the residual because the $t$th observation was not used in estimating the value of $\hat{y_t}$.)

2. Repeat step 1 for $t$ = 1,…,T.
3. Compute the MSE from $e_1^*, ..., e_T^*$. We shall call this the <b>CV</b>

Although this looks like a time-consuming procedure, there are fast methods of calculating CV, so that it takes no longer than fitting one model to the full data set. The equation for computing CV efficiently is given in Section 7.9. Under this criterion, the best model is the one with the smallest value of CV.

## Akaike's Information Criterion (lower is better)

A closely-related method is Akaike’s Information Criterion, which we define as:

$AIC = T\text{log}(\frac{SSE}{T}) + 2(k + 2)$

where T is the number of observations used for estimation and k is the number of predictors in the model. Different computer packages use slightly different definitions for the AIC, although they should all lead to the same model being selected. The k+2 part of the equation occurs because there are k+2 parameters in the model: the k coefficients for the predictors, the intercept and the variance of the residuals. The idea here is to penalise the fit of the model (SSE) with the number of parameters that need to be estimated.

## Corrected Akaike's Information Criterion (lower is better)

For small values of T, the AIC tends to select too many predictors, and so a bias-corrected version of the AIC has been developed:

$AIC_c = AIC + \frac{2(k+2)(k+3)}{T - k - 3}$

## Schwarz's Bayesian Information Criterion

A related measure is Schwarz’s Bayesian Information Criterion

$BIC = T\text{log}(\frac{SSE}{T}) + (k + 2)log(T)$

## Which measure should we use?

While R^2^ is widely used, and has been around longer than the other measures, its tendency to select too many predictor variables makes it less suitable for forecasting.

Many statisticians like to use the BIC because it has the feature that if there is a true underlying model, the BIC will select that model given enough data. However, in reality, there is rarely, if ever, a true underlying model, and even if there was a true underlying model, selecting that model will not necessarily give the best forecasts (because the parameter estimates may not be accurate).

Consequently, we recommend that one of the AICc, AIC, or CV statistics be used, each of which has forecasting as their objective. If the value of T
is large enough, they will all lead to the same model. In most of the examples in this book, we use the AICc value to select the forecasting model.

## Example: US consumption

In the multiple regression example for forecasting US consumption we considered four predictors. With four predictors, there are 2^4^=16 possible models. Now we can check if all four predictors are actually useful, or whether we can drop one or more of them. All 16 models were fitted and the results are summarised in Table 7.1. A X indicates that the predictor was included in the model, Hence the first row shows the measures of predictive accuracy for a model including all four predictors.

The results have been sorted according to the AICc. Therefore the best models are given at the top of the table, and the worst at the bottom of the table.
Table 7.1: All 16 possible models for forecasting US consumption with 4 predictors.

| Income| Production | Savings | Unemployment | Adj R^2^ | CV | AIC | AICc | BIC |
|:-----:|:----------:|:-------:|:------------:|:--------:|:--:|:---:|:----:|:---:|
|X|X|X|X|0.763|0.104|-456.6|-456.1|-436.9|
|X|X|X||0.761|0.105|-455.2|	-454.9|-438.7|
|X||X|X|0.760|0.104|-454.4|-454.1|-437.9|
|X||X||0.735 |	0.114 |	-435.7 |	-435.5 |	-422.6|
|X|X| |X|0.366 |	0.271 |	-262.3 |	-262.0 |	-245.8|
||X|X|X|0.349 |	0.279 |	-257.1 |	-256.8 |	-240.7|
|X|||X|0.345 |	0.276 |	-256.9 |	-256.6 |	-243.7|
|X|X|||0.336 |	0.282 |	-254.2 |	-254.0 |	-241.0 |
||X|X||0.324 |	0.287 |	-250.7 |	-250.5 |	-237.5 |
|||X|X|0.311 |	0.291 |	-246.9 |	-246.7 |	-233.7 |
||X||X|0.308 |	0.293 |	-246.1 |	-245.9 |	-232.9|
||X|||0.276 |	0.304 |	-238.1 |	-238.0 |	-228.2 |
||||X|0.274 |	0.303 |	-237.4 |	-237.3 |	-227.5 |
|X||||0.143 |	0.356 |	-204.6 |	-204.5 |	-194.7|
||X|||0.061 |	0.388 |	-186.5 |	-186.4 |	-176.7 |
|||||0.000 |	0.409 |	-175.1 |	-175.0 |	-168.5 |

## Best subset regression

### Stepwise regression

If there are a large number of predictors, it is not possible to fit all possible models. For example, 40 predictors leads to ^40^ > 1 trillion possible models! Consequently, a strategy is required to limit the number of models to be explored.

An approach that works quite well is backwards stepwise regression:

•    Start with the model containing all potential predictors.
•    Remove one predictor at a time. Keep the model if it improves the measure of predictive accuracy.
•    Iterate until no further improvement.

If the number of potential predictors is too large, then the backwards stepwise regression will not work and <i>forward stepwise regression<i> can be used instead. This procedure starts with a model that includes only the intercept. Predictors are added one at a time, and the one that most improves the measure of predictive accuracy is retained in the model. The procedure is repeated until no further improvement can be achieved.

### Beware of inference <b>after</b> selecting predictors

We do not discuss statistical inference of the predictors in this book (e.g., looking at p-values associated with each predictor). If you do wish to look at the statistical significance of the predictors, beware that <mark>any procedure involving selecting predictors first will invalidate the assumptions behind the p-values. The procedures we recommend for selecting predictors are helpful when the model is used for forecasting; they are not helpful if you wish to study the effect of any predictor on the forecast variable.</mark>

Completion through section 7.5