---
title: "9. ARIMA models.RMD"
author: "Russ Conte"
date: "10/30/2021"
output: html_document
---
# 9. ARIMA models

ARIMA models provide another approach to time series forecasting. Exponential smoothing and ARIMA models are the two most widely used approaches to time series forecasting, and provide complementary approaches to the problem. While exponential smoothing models are based on a description of the trend and seasonality in the data, ARIMA models aim to describe the autocorrelations in the data.

Before we introduce ARIMA models, we must first discuss the concept of stationarity and the technique of differencing time series.


## 9.1 Stationarity and differencing

A stationary time series is one whose statistical properties do not depend on the time at which the series is observed.15 Thus, time series with trends, or with seasonality, are not stationary — the trend and seasonality will affect the value of the time series at different times. On the other hand, a white noise series is stationary — it does not matter when you observe it, it should look much the same at any point in time.

In general, a stationary time series will have no predictable patterns in the long-term. Time plots will show the series to be roughly horizontal (although some cyclic behaviour is possible), with constant variance.

## Differencing

Transformations such as logarithms can help to stabilise the variance of a time series. Differencing can help stabilise the mean of a time series by removing changes in the level of a time series, and therefore eliminating (or reducing) trend and seasonality.

As well as the time plot of the data, the ACF plot is also useful for identifying non-stationary time series. <mark>For a stationary time series, the ACF will drop to zero relatively quickly, while the ACF of non-stationary data decreases slowly. Also, for non-stationary data, the value of $r_1$ is often large and positive.</mark>

```{r Figure 9.2: The ACF of the Google closing stock price in 2015 (left) and of the daily changes in Google closing stock price in 2015 (right).}

library(tidyverse)
library(fpp3)

google_stock <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2015) %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE)
# Filter the year of interest
google_2015 <- google_stock %>% filter(year(Date) == 2015)

google_2015 %>% 
  mutate(diff_close = difference(Close)) %>% 
  features(diff_close, ljung_box, lag = 10)

```

Let's look at an autocorrelation of the Close price of Goolge 2015 stock:

```{r Autocorrelation Function (ACF) of Google 2015 stock}

google_2015 %>% 
  ACF(Close) %>% 
  autoplot()

```

Let's look at a graph of the lag of the Close price of Google2015

```{r the daily changes in Google closing stock price in 2015}
google_2015 %>% 
  mutate(diff_close = difference(Close)) %>% 
  ACF(diff_close) %>% 
  autoplot()

```

The ACF of the differenced Google stock price looks just like that of a white noise series. <mark>Only one autocorrelation is outside of the 95% limits,</mark> and the Ljung-Box $Q^*$statistic has a p-value of 0.637 (for $h =$10). This suggests that the daily change in the Google stock price is essentially a random amount which is uncorrelated with that of previous days.

## Random walk model

The differenced series is the <i>change</i> between consecutive observations in the original series, and can be written as:

$y^{'}_t = y_t - y_{t-1}$<br>

The differenced series will have only $T - 1$ values, since it is not possible to calculate a difference, $y^{'}$ for the first observation.

When the differenced series is white noise, the model for the original series can be written as:

$y_t - y_{t-1} = \epsilon_t$<br>

where $\epsilon_t$ denotes white noise. Rearranging this leads to the "random walk" model,

$y_t = y_{t-1} + \epsilon_t$

Random walk models are widely used for non-stationary data, particularly financial and economic data. Random walks typically have:

• Long periiods of apprent trneds up or down
• Sudden and unpredictable changes in direction

The forecasts from a random walk model are equal to the last observation, as future movements are unpredictable, and are equally likely to be up or down. Thus, the random walk model underpins naïve forecasts, first introduced in Section 5.2.

A closely related model allows the differences to hvaswe a non-zero mean. Then:

$y_t - y_{t-1} = c + \epsilon_t$ or $y_t = c + y_{t-1} + \epsilon_t$

The value of $c$ is the average of the changes between consecutive observations. If $c$ is positive, then the average change is an increase in the value of $y_t$. Thus, $y_t$ will tend to drift upwards. However, if $c$ is negative, $y_t$ will tend to drift downwards.

This is the model behind the drift method, also discussed in Section 5.2.

## Second-order differencing

Occasionally the differenced data will not appear to be stationary and it may be necessaryto difference the data a second time to obtain a stationary series:

$y_t^{''} = y_t^{'} - y_{t-1}^{'}$

## Seasonal differencing

A seasonal difference is the difference between an observation and the previous observation from the same season. So:

$y_t^{'} = y_t - y_{t-m},$

where $m$= the number of seasons.These are also called "lag-m differences," as we subtract the observation after a lag of $m$ periods.

If the seasonaly differenced data appear to ber white noise, then an appropriate model for the original data is:

$y_t = y_{t-m} + \epsilon_t$<br>

Forecasts from this model are equal to the last observation from the relevant season. That is, <marl>this model gives seasonal näive forecasts, introduced in Section 5.2</mark>

To distinguish seasonal differences from ordinary differences, we sometimes refer to ordinary differences as “first differences,” meaning differences at lag 1.

Sometimes it is necessary to take both a seasonal difference and a first difference to obtain stationary data. Figure 9.4 plots Australian corticosteroid drug sales ($AUD) (top panel). Here, the data are first transformed using logarithms (second panel), then seasonal differences are calculated (third panel). The data still seem somewhat non-stationary, and so a further lot of first differences are computed (bottom panel).

```{r Figure 9.4: Top panel: Corticosteroid drug sales ($AUD).}

PBS %>% 
  filter(ATC2 == "H02") %>% 
  summarise(Cost = sum(Cost)/1e6) %>% 
  transmute(
    `Sales ($million)` = Cost,
    `Log sales` = log(Cost),
    `Annual change in log sales` = difference(log(Cost), 12),
    `Doubly differenced log sales` = difference(difference(log(Cost), 12), 1)
    ) %>% 
  pivot_longer(-Month, names_to = "Type", values_to = "Sales") %>% 
  mutate(
    Type = factor(Type, levels = c(
      "Sales ($million)",
      "Log sales",
      "Annual change in log sales",
      "Doubly differenced log sales"))
  ) %>% 
  ggplot(aes(x = Month, y = Sales)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Corticosteroid drug sales", y = NULL)

```

There is a degree of subjectivity in selecting which differences to apply. The seasonally differenced data in Figure 9.3 do not show substantially different behaviour from the seasonally differenced data in Figure 9.4. In the latter case, we could have decided to stop with the seasonally differenced data, and not done an extra round of differencing. In the former case, we could have decided that the data were not sufficiently stationary and taken an extra round of differencing. Some formal tests for differencing are discussed below, but there are always some choices to be made in the modelling process, and different analysts may make different choices.

If $y_t^{'} =y_t - y_{t-m}$ denotes a seasonally differenced series, then the twice-differenced series is:

$y_t^{''} = y_t^{'} - y_{t-1}^{'}$
$= (y_t - y_{t-m}) - (t_{t-1} - y_{t - m - 1})$
$= y_t - t_{t-1} - y_{t-m} + y_{t-m-1}$

When both seasonal and first differences are applied, it makes no difference which is done first—the result will be the same. <mark>However, if the data have a strong seasonal pattern, we recommend that seasonal differencing be done first, because the resulting series will sometimes be stationary and there will be no need for a further first difference.</mark> If first differencing is done first, there will still be seasonality present.

Beware that applying more differences than required will induce false dynamics or autocorrelations that do not really exist in the time series. Therefore, <mark>do as few differences as necessary to obtain a stationary series.</mark>

It is important that if differencing is used, the differences are interpretable. First differences are the change between one observation and the next. Seasonal differences are the change between one year to the next. Other lags are unlikely to make much interpretable sense and should be avoided.

## Unit root tests

One way to determine more objectively whether differencing is required is to use a <i>unit root test</i>. These are statistical hypothesis tests of stationarity that are designed for determining whether differencing is required.

Let's apply the <i>Kwiatkowski-Phillips-Schmidt-Snit (KPSS)</i> test to the Goolge 2015 stock price data:

```{r Test the google 2015 stock price to test if differencing is required}
google_2015 %>% 
  features(Close, unitroot_kpss)

```

The p-value is reported as 0.01 if it is less than 0.01, and as 0.1 if it is greater than 0.1. In this case, the test statistic (3.56) is bigger than the 1% critical value, so the p-value is less than 0.01, indicating that the null hypothesis is rejected. That is, the data are not stationary. We can difference the data, and apply the test again.

```{r}
google_2015 %>% 
  mutate(diff_close = difference(Close)) %>% 
  features(diff_close, unitroot_kpss)
```

his time, the test statistic is tiny, and well within the range we would expect for stationary data, so the p-value is greater than 0.1. We can conclude that the differenced data appear stationary.

This process of using a sequence of KPSS tests to determine the appropriate number of first differences is carried out using the unitroot_ndiffs() feature.

```{r}
google_2015 %>% 
  features(Close, unitroot_ndiffs)
```

As we saw from the KPSS tests above, one difference is required to make the `google_2015` data stationary.

A similar feature for determining whether seasonal differencing is required is `unitroot_nsdiffs()`, which uses the measure of seasonal strength introduced in Section 4.3 to determine the appropriate number of seasonal differences required. No seasonal differences are suggested if  $F_s \lt 0.64$, otherwise one seasonal difference is suggested.

We can apply the `unitroot_nsdiffs()` to the monthly total Australian retail turnover.

```{r}
aus_total_retail <- aus_retail %>% 
  summarise(Turnover = sum(Turnover))

aus_total_retail %>% 
  mutate(log_turnover = log(Turnover)) %>% 
  features(log_turnover, unitroot_nsdiffs)

```


```{r}
aus_total_retail %>% 
  mutate(log_turnover = difference(log(Turnover), 12)) %>% 
  features(log_turnover, unitroot_nsdiffs)
```

Because unitroot_nsdiffs() returns 1 (indicating one seasonal difference is required), we apply the unitroot_ndiffs() function to the seasonally differenced data. These functions suggest we should do both a seasonal difference and a first difference.

## 9.2 Backshift notation

The backward shift operator $B$ is a useful notationla device when working with time series lags:

$B_{y_{t}} = y_{t-1}$

In other words, $B$, operating on $y_t$, has the effect of shifting the data back one period. Two applications of $B$ to $y_t$ shifts the data back two periods:

$B(B_{y_{t}}) = B^2y_t = y_{t-2}$

For monthly data, if we wish to consider "the same month last year," the notation is $B^{12}y_t = y_{t -12}$

The backward shift operator is convenient for describing the process of <i>differencing</i>. The first difference can be written as:

$y_t^{'} = y_t - y_{t-1} = y_t - By_t = (1 - B)y_t$

So the first difference can be represented by $(1 - B)$. Similary, if second-order differences have to be computed, then:

$y_t^{''} = y_t - 2y_{t-1} + y_{t-2} = (1 - 2B + B^2)y_t = (1 - B)^2y_t$

In general, the dth-order difference can be written as:

$(1 - B)^d y_t$

Backshift notation is particularly useful when combining differences, as the operator can be treated using ordinary algebraic rules. In particular, terms involving $B$ can be multiplied together.

In a multiple regression model, introduced in Chapter 7, we forecast the variable of interest using a linear combination of predictors. In an autoregression model, we forecast the variable of interest using a linear combination of <i>past values of the variable</i>. The term autoregression indicates that it is a regression of the variable against itself.

Thus, an autoregressive model of order $p$ can be written as:

$y_t = c + \phi_1y_{t-1} + \phi_2y_{t-2} + ... + \phi_py_{t-p} + \epsilon_t$

where $\epsilon_t$ is white noise. This is like a multiple regression but with lagged values of $y_t$ as predictors. We refer to this as an <b>AR($p$)</b> model, an autoregressive model of order $p$.

Autoregressive models are remarkably flexible at handling a wide range of different time series patterns. The two series in Figure 9.5 show series from an AR(1) model and an AR(2) model. Changing the parameters $\phi_1,...\phi_p$esults in different time series patterns. The variance of the error term $\epsilon_t$ will only change the scale of the series, not the patterns.

We normally restrict autoregressive models to stationary data, in which case some constraints on the values of the parameters are required:

• For an AR(1) model: $-1 \lt \phi_1 \lt 1$<br>
• For an AR(2) model: $-1 \lt \phi_2 \lt1, \phi_1 +\phi_2 \lt 1, \phi_2 - \phi_1 <1$

When $p \geq3$, the restrictions are much more complicated. The `fable` package takes care of these restrictions when estimating a model.

## 9.4 The moving average models

Rather than using past values of the forecast variable in a regression, a moving average model uses past forecast errors in a regression-like model,

$y_t = c + \epsilon_t + \theta_1 \epsilon_{t-1} +\theta_2 \epsilon_{t-2} +\theta_q \epsilon_{t-q}$

where $\epsilon_t$ is white noise. We refer to this as an <b>MA(q) model</b>, a moving average model of order $q$. Of course, we do not <i>observe</i> the value of $\epsilon_t$, so it is not really a regression in the usual sense.

Notice that each value of $y_t$ can be thought of as a weighted moving average of the past few forecast errors (although the coefficients will not normally sum to one). However, moving average models should not be confused with the moving average smoothing we discussed in Chapter 3. A moving average model is used for forecasting future values, while moving average smoothing is used for estimating the trend-cycle of past values.

Figure 9.6 shows some data from an MA(1) model and an MA(2) model. Changing the parameters $\theta_1,...,\theta_q$ results in different time series patterns. As with autoregressive models, the variance of the error term $\epsilon_t$ will only change the scale of the series, not the patterns.

It is possible to write any stationary AR($p$) model as an MA($\infty$) model. For example, using repeated substitution, we can demonstrate this for an AR(1) model:

$y_t = \phi y_{t-1} + \epsilon_t$<br>
$\phi_1(\phi_1 y_{t-2} + \epsilon_{t-1}) + \epsilon_t$<br>
$\phi_1^2 y_{t-2} + \phi \epsilon_{t-1} + \epsilon_t$<br>
$\phi_1^3 y_{t-3} + \phi^2 \epsilon_{t-2} + \phi_1\epsilon_{t-1} + \epsilon_t$<br>
etc.

Provided $-1 \lt \phi_1 \lt 1$ the value of $\phi_1^k$ will get smaller as $k$ gets larger. So eventually we obtain:

$y_t = \epsilon_t + \phi_1 \epsilon_{t-1} + \phi_1^2 \epsilon_{t-2} + \phi_1^3 \epsilon_{t-3} + ...,$

am MA($\infty$) process.

The reverse result holds if we impose some constraints on the MA parameters. Then the MA model is called invertible. That is, we can write any invertible MA($q$) process as an AR($\infty$) process. Invertible models are not simply introduced to enable us to convert from MA models to AR models. They also have some desirable mathematical properties.

For example, consider the MA(1) process, $y_t = \epsilon_t + \theta_1 \epsilon_{t-1}$. In its AR($\infty$) representation, the most recent error can be written as a linear function of current and past observations:

$\epsilon_t = \sum_{j=0}^\infty (-\theta_1)^j y_{t-j}$

When $\left| \theta_1 \right|\gt 1$, he weights increase as lags increase, so the more distant the observations the greater their influence on the current error. When $\left| \theta_1 \right| = 1$ the weights are constant in size, and the distant observations have the same influence as the recent observations. As neither of these situations make much sense, we require $\left| \theta_1 \right| \lt 1$ so the most recent observations have higher weight than observations from the more distant past. Thus, the process is invertible when $\left| \theta_1 \right| \lt 1$.

The invertaibility constraints for other models are similar to the stationarity constraints.

•For an MA(1) model: $-1 \lt \theta_1 \lt 1$<br>
•For an MA(2) model: $-1 \lt \theta_2 \lt 1, \theta_2 + \theta_1 \gt -1, \theta_1 - \theta_2 \lt 1$<br>

More complicated conditions hold for $q \ge 3$. Again, the `fable` package will take care of these constraints when estimating the models.

# 9.5 Non-seasonal ARIMA models

If we combine differencing with autoregression and a moving average model, we obtain a non-seasonal ARIMA model. ARIMA is an acronym for AutoRegressive Integrated Moving Average (in this context, “integration” is the reverse of differencing). The full model can be written as

$y_t^{'} = c + \phi_1y_{t-1}^{'}+...+\phi_py_{t-p}^{'} + \theta_1\epsilon_{t-1} + ... +\theta_q\epsilon_{t-q} + \epsilon_t$<br>

where $y'_{t}$ is the differenced series (it may have been differenced more than once). The “predictors” on the right hand side include both lagged values of $y_t$ and lagged errors. We cal this an <b>ARIMA(p, d, q) model</b> where

||
|:-----:|
|$p$ = the order of the autoregressive part|
|$d$ = degree of the first differencing involved|
|$p$ = order of the moving average part|

The same stationarity and invertibility conditions that are used for autoregressive and moving average models also apply to an ARIMA model.

Many of the models we have already discussed are special cases of the ARIMA model, as shown in Table 9.1.

<center>Table 9.1: Special cases of ARIMA models.</center>
|||
|:-----:|:-----:|
|White noise|ARIMA(0,0,0) with no constant|
|Random walk|ARIMA(0,1,0) with no constant|
|Random walk with drift|ARIMA(0,1,0) with a constant|
|Autoregression|ARIMA($p$,0,0)|
|Moving average|ARIMA(0,0,$q$)|

Once we start combining components in this way to form more complicated models, it is much easier to work with the backshift notation. For example, Equation (9.1) can be written in backshift notation as

$(1 - \phi_1B - \cdots - \phi_pB^p) \quad (1-B)^d y_t \quad= \quad c +(1 + \theta_1B + \cdots + \theta_qB^q) \epsilon_t$

$\qquad \qquad \uparrow \qquad \qquad \qquad  \quad \uparrow \qquad \qquad \qquad \quad \qquad \uparrow$

$\qquad \quad AR(p) \qquad \qquad \text{d-differences} \qquad \qquad \qquad MA(q)$

